\bf Climbing

### 2021-07-26 manuscript/Rmd/draft.tex
Create Rmd manuscript folder

---
title: Climbing

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Author 1
  thanks: 
  affiliation:

- name: Author 2
  affiliation:

- name: Author 3
  affiliation:

keywords:
- sport climbing
- scoring system

abstract: |
  This manuscript... The main results are... 
# bibliography: bib_example.bib
output: rticles::asa_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "",
                      fig.width = 5,
                      fig.height = 4,
                      results = "asis")
options(xtable.comment = FALSE,
        xtable.caption.placement = "top", 
        xtable.include.rownames = FALSE)
```

# Introduction

@@

# Data and Methods

# Results

## Simulations: Uniform Ranks

```{r}
library(tidyverse)
library(knitr)
library(xtable)
theme_set(theme_bw())
```

```{r}
set.seed(1)
nsim <- 10000
nplay <- 20
qual_sim <- list()
for (i in 1:nsim) {
  qual_sim[[i]] <-
    bind_cols(
      player = 1:nplay,
      e1 = sample(1:nplay, replace = FALSE),
      e2 = sample(1:nplay, replace = FALSE),
      e3 = sample(1:nplay, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

qual <- bind_rows(qual_sim) %>% 
  mutate(qual_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  mutate(qual_rank = rank(qual_score, ties.method = "random")) %>%
  ungroup()
```

```{r}
set.seed(1)
nsim <- 10000
nplay <- 8
final_sim <- list()
for (i in 1:nsim) {
  final_sim[[i]] <-
    bind_cols(
      player = 1:nplay,
      e1 = sample(1:nplay, replace = FALSE),
      e2 = sample(1:nplay, replace = FALSE),
      e3 = sample(1:nplay, replace = FALSE)
    ) %>%
    mutate(sim = i)
}

final <- bind_rows(final_sim)

final <- final %>% 
  mutate(final_score = e1 * e2 * e3) %>% 
  group_by(sim) %>% 
  mutate(final_rank = rank(final_score, ties.method = "random")) %>%
  ungroup()
```

@@

```{r, fig.cap = "Distribution of qualification scores"}
qual %>%
  ggplot(aes(qual_score)) +
  geom_histogram(bins = 20, color = "white") + 
  labs(x = "Qualification Score",
       y = "Frequency")
  # theme(axis.title = element_text(size = 9),
  #       axis.text = element_text(size = 8))
```

For the qualification round, a climber is almost guaranteed to make the final round if they win the first event (with a 99.51% chance of advancing) or if they win at least one of the three climbing concentrations (99.48%). On the other hand, finishing last in the first event or in any event would certainly hurt an athlete's chance of finishing in the top 8, as the probabilities of a climber advancing given they finish last in the first and in any event are 0.1830 and 0.1885, respectively. In addition, the average score for qualification positions 1 to 8 are displayed in Table 1. We notice that on average, the minimum score that one should aim for in order to move on to the final round is 435 (for 8th rank).

```{r}
qual %>% 
  group_by(qual_rank) %>% 
  summarize(avg_adv_score = mean(qual_score)) %>% 
  filter(qual_rank <= 8) %>% 
  xtable(caption = "Average score for each qualifying rank") %>% 
  print()
```

Regarding the finals, a climber is very likely to earn a medal (finish in the top 3) if they win the first event (83.03% chance) or any event (85.01%). In order to obtain a climbing medal, the average score for getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648, respectively (Table 2). A notable trend we observe for both qualification and final rounds is as the rank increases, the distribution of the scores becomes more spread out. (Figure..., facet)

```{r}
final %>% 
  group_by(final_rank) %>% 
  summarize(avg_score = mean(final_score)) %>% 
  xtable(caption = "Average score of medalists") %>% 
  print()
```

```{r, fig.cap = "Boxplots of scoring distribution for every qualification rank"}
qual %>% 
  mutate(qual_rank = factor(qual_rank)) %>% 
  ggplot(aes(x = qual_score, y = qual_rank)) +
  geom_boxplot()
```

## Drawbacks of the scoring system

### Why product of ranks instead of sum

@@

### Speed climbing vs lead and bouldering

We collected data on major climbing competitions from 2018 to 2020, including the 2020 Continental Championships of Europe, Africa, Oceania, Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and 2018 Asian Games. We are interested in looking at the relationships between the ranks of the individual events and the final standings, and we computed Kendall’s Tau (Kendall Rank Correlation Coefficient).

@@

```{r}
wq <- read_csv("https://raw.githubusercontent.com/qntkhvn/climbing/main/data/2018_youth_olympics/women_qual.csv")
library(GGally)
```

```{r, fig.cap = "Kendall's rank correlations - 2018 World Championship, Women's Qualification"}
wq %>% 
  select(rank, speed, bould, lead) %>% 
  ggpairs(diag = "blank",
          upper = list(continuous = wrap("cor", method = "kendall")))
```

```{r}
# wq %>% 
#   select(rank, speed:lead) %>% 
#   cor(method = "kendall") %>% 
#   data.frame()
```

### Drop and re-rank

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

%% load any required packages here

% Pandoc citation processing

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Climbing}

  \author{
        Author 1 \\
    \\
     and \\     Author 2 \\
    \\
     and \\     Author 3 \\
    \\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Climbing}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
This manuscript\ldots{} The main results are\ldots{}
\end{abstract}

\noindent%
{\it Keywords:} sport climbing, scoring system
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{data-and-methods}{%
\section{Data and Methods}\label{data-and-methods}}

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{simulations-uniform-ranks}{%
\subsection{Simulations: Uniform
Ranks}\label{simulations-uniform-ranks}}

@@

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{Distribution of qualification scores}
\end{figure}

For the qualification round, a climber is almost guaranteed to make the
final round if they win the first event (with a 99.51\% chance of
advancing) or if they win at least one of the three climbing
concentrations (99.48\%). On the other hand, finishing last in the first
event or in any event would certainly hurt an athlete's chance of
finishing in the top 8, as the probabilities of a climber advancing
given they finish last in the first and in any event are 0.1830 and
0.1885, respectively. In addition, the average score for qualification
positions 1 to 8 are displayed in Table 1. We notice that on average,
the minimum score that one should aim for in order to move on to the
final round is 435 (for 8th rank).

\begin{table}[ht]
\centering
\caption{Average score for each qualifying rank} 
\begin{tabular}{rr}
  \hline
qual\_rank & avg\_adv\_score \\ 
  \hline
  1 & 36.02 \\ 
    2 & 73.61 \\ 
    3 & 115.40 \\ 
    4 & 162.23 \\ 
    5 & 216.00 \\ 
    6 & 278.16 \\ 
    7 & 350.33 \\ 
    8 & 434.59 \\ 
   \hline
\end{tabular}
\end{table}

Regarding the finals, a climber is very likely to earn a medal (finish
in the top 3) if they win the first event (83.03\% chance) or any event
(85.01\%). In order to obtain a climbing medal, the average score for
getting gold, silver, and bronze are 9.6748, 20.4143, and 33.2648,
respectively (Table 2). A notable trend we observe for both
qualification and final rounds is as the rank increases, the
distribution of the scores becomes more spread out. (Figure\ldots,
facet)

\begin{table}[ht]
\centering
\caption{Average score of medalists} 
\begin{tabular}{rr}
  \hline
final\_rank & avg\_score \\ 
  \hline
  1 & 9.67 \\ 
    2 & 20.41 \\ 
    3 & 33.26 \\ 
    4 & 50.59 \\ 
    5 & 74.76 \\ 
    6 & 110.05 \\ 
    7 & 164.43 \\ 
    8 & 265.78 \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{Boxplots of scoring distribution for every qualification rank}
\end{figure}

\hypertarget{drawbacks-of-the-scoring-system}{%
\subsection{Drawbacks of the scoring
system}\label{drawbacks-of-the-scoring-system}}

\hypertarget{why-product-of-ranks-instead-of-sum}{%
\subsubsection{Why product of ranks instead of
sum}\label{why-product-of-ranks-instead-of-sum}}

@@

\hypertarget{speed-climbing-vs-lead-and-bouldering}{%
\subsubsection{Speed climbing vs lead and
bouldering}\label{speed-climbing-vs-lead-and-bouldering}}

We collected data on major climbing competitions from 2018 to 2020,
including the 2020 Continental Championships of Europe, Africa, Oceania,
Pan-America; 2019 and 2018 World Championships; 2018 Asian Games; and
2018 Asian Games. We are interested in looking at the relationships
between the ranks of the individual events and the final standings, and
we computed Kendall's Tau (Kendall Rank Correlation Coefficient).

@@

\begin{figure}
\centering
\includegraphics{draft_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{Kendall's rank correlations - 2018 World Championship, Women's
Qualification}
\end{figure}

\hypertarget{drop-and-re-rank}{%
\subsubsection{Drop and re-rank}\label{drop-and-re-rank}}

\bibliographystyle{agsm}
\bibliography{}

\end{document}
